### 2019-04-09 Tex/elsevier/list of abbreviations.tex
new text

list of abbreviations, units and nomenclature

$a$ Ideality or quality factor
AC Alternating current
$C_{10}$ 10h of rated battery capacity
$C_{bank}$ Total capacity of the battery bank
CBMC C Bounded Model Checker
$ cos \varphi $ Power factor of the inverter
CPA Configurable Program Analysis
CPAchecker Configurable Program Analysis Checker
DC Direct current
$DOD$ Depth of discharge ($DOC=1-SOC$)
$E_{consumption}$ Energy consumption
$E_{corrected}$ Energy corrected
$E_{p}$ Energy produced by one panel
ESBMC Efficient SMT-based Bounded Model Checker
$G$ Solar irradiance
$ G_{ref} $ Solar irradiance at STC ($1000 W/m^{2}$)
$I$ Current
$I_{0}$ Reverse saturation current
$ I_{AC} $ Current that the inverter supply to the house
$I_{c,min}$ Charge controller minimum current
$ I_{DC} $ Current required by the inverter from the DC source 
$ I_{in} $ Input current
$I_{load}$ Current that feed the load
$ I_{out} $ Output current
$I_{ph}$ Photocurrent
$ I_{ph,ref} $ Photocurrent at STC
$I_{pv}$ Current from the PV system
$ I_{sc} $ Short-circuit current
$ I_{sc,amb} $ Ambient short-circuit current
$ I_{sc,ref} $ Short-circuit current at STC
$ k_{B} $ Boltzmann constant ($ 1.3806503\times10^{-23}J/K $); 
$MAX_{AC,ref}$ Peak power that the inverter can support
$N$ Number of series-connected cells
$N_{B}total$ Sized number of batteries
$N_{BP}$ Sized number of parallel batteries
$N_{BP}min$ Total minimum number of parallel batteries
$N_{BS}$ Sized number of series batteries
$N_{BS}min$ Total minimum number of series batteries
NOCT Nominal Operating Cell Temperature
$N_{PS}$ Sized number of panels in series
$N_{PSmin}$ Total number of panels in series
$N_{PP}$ Sized number of panels in parallel
$N_{PPmin}$ Total umber of panels in parallel
$ N_{TP} $ Sized number of panels
$N_{TPmin}$ Total minimum number of needed solar panels
PV Photovoltaic
P Power
$P_{in}$ Input power
$P_{out}$ Output power
$P_{surge}$ Maximum peak power demanded by house for few seconds
$q$ Absolute value of the electron's charge ($ -1.60217646\times10^{-19}C $)
SAT Satisfiability or Boolean satisfiability problem 
SAT When returns a failure from the verification tool
SMT Satisfiability modulo theories
$ SOC $ Battery state of charge
STC Standard Test Condition (reference state)
$T$ Temperature
$ T_{air} $ Ambient temperature
$ T_{ref}$ Temperature at STC ($=298.15 K$ or $25^{o}C$)
UNSAT Condition when verification tool succeeds (could not reach an error)
V Voltage
$ V_{AC}  $ Voltage that the inverter supply to the house
$V_{c}$ Battery charge voltage 
$V_{d}$ Battery discharge voltage
$ V_{DC} $ Input voltage to the inverter delivered by the DC source
$ V_{in} $ Input voltage
$ V_{oc} $ Open-circuit voltage
$ V_{oc,ref} $ Reference open-circuit voltage at STC
$ V_{out} $ Output voltage
$ V_{system} $ DC voltage of the bus
$V_{T}$ Thermal voltage ($ V_{T}=k_{B}T/q $)
$\Delta T$ Temperature variation ($\Delta T=T-T_{ref} $)
$ \mu_{I} $ Short-circuit current temperature coefficient ($A/K$)
$ \mu_{V} $ Open-circuit voltage temperature coefficient ($ V/K $)
$\eta_{b}$ Battery efficiency
$\eta_{c}$ Charge controller efficiency
$\eta_{i}$ Inverter efficiency
$\eta_{p}$ PV panel efficiency