### 2019-06-27 Tex/Old text/chapters/chapter03.tex
thesis structure

chapter03 here amamas asd asd as das das d af dsgf dfg  tre tre t ret ret er ter te rt ert ert er tre t