### 2019-06-27 Tex/Old text/main.tex

\documentclass[12pt,twoside]{report}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage{caption}
\usepackage{subcaption}
\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm,bindingoffset=6mm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[RO,LE]{Thesis Title}
\fancyfoot{}
\fancyfoot[LE,RO]{\thepage}
\fancyfoot[LO,CE]{Chapter \thechapter}
\fancyfoot[CO,RE]{Author Name}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

%\usepackage[style=authoryear,sorting=none]{biblatex}
%\addbibresource{references.bib}

\title{Thesis Title}
\author{Alessandro Trindade}
\date{Day Month Year}

\begin{document}

\input{titlepage}

\input{abstract}

\chapter*{Dedication}
To mum and dad

\chapter*{Declaration}
I declare that..

\chapter*{Acknowledgements}
I want to thank...

\tableofcontents

\listoffigures

\listoftables

\chapter{Introduction}
\input{chapters/introduction}

\chapter{Chapter Two Title}
\input{chapters/chapter02}

\chapter{Chapter Three Title}
\input{chapters/chapter03}

\chapter{Chapter Four Title}
\input{chapters/chapter04}

\chapter{Conclusion}
\input{chapters/conclusion}

\appendix
\chapter{Appendix Title}
\input{chapters/appendix}

%\printbibliography
\bibliography{trindadeThesis}{}

\end{document}