### 2019-07-02 Tex/Old text/fichacatalog.tex
thesis updated text

\thispagestyle{plain}
\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
    \vspace{0.9cm}
    \textbf{Ficha Catalográfica}
\end{center}
\thispagestyle{plain}
Ficha catalog aqui