### 2019-07-03 Tex/Old text/abbreviations.tex
thesis updated text

\thispagestyle{plain}
\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
    \vspace{0.9cm}
    \textbf{List of Abbreviations, Units and NomenclatureList of Abbreviations, Units and Nomenclature}
\end{center}

\noindent $a$ Ideality or quality factor \\
AC Alternating current \\
$C_{10}$ 10h of rated battery capacity \\
$C_{bank}$ Total capacity of the battery bank \\
CBMC C Bounded Model Checker \\
$ cos \varphi $ Power factor of the inverter \\
CPA Configurable Program Analysis \\
CPAchecker Configurable Program Analysis Checker \\
DC Direct current \\
$DOD$ Depth of discharge ($DOC=1-SOC$) \\
$E_{consumption}$ Energy consumption \\
$E_{corrected}$ Energy corrected \\
$E_{p}$ Energy produced by one panel \\
ESBMC Efficient SMT-based Bounded Model Checker \\
$G$ Solar irradiance \\
$ G_{ref} $ Solar irradiance at STC ($1000 W/m^{2}$) \\
$I$ Current \\
$I_{0}$ Reverse saturation current \\
$ I_{AC} $ Current that the inverter supply to the house \\
$I_{c,min}$ Charge controller minimum current \\
$ I_{DC} $ Current required by the inverter from the DC source \\
$ I_{in} $ Input current \\
$I_{load}$ Current that feed the load \\
$ I_{out} $ Output current \\
$I_{ph}$ Photocurrent \\
$ I_{ph,ref} $ Photocurrent at STC \\
$I_{pv}$ Current from the PV system \\
$ I_{sc} $ Short-circuit current \\
$ I_{sc,amb} $ Ambient short-circuit current \\
$ I_{sc,ref} $ Short-circuit current at STC \\
$ k_{B} $ Boltzmann constant ($ 1.3806503\times10^{-23}J/K $); \\
$MAX_{AC,ref}$ Peak power that the inverter can support \\
$N$ Number of series-connected cells \\
$N_{B}total$ Sized number of batteries \\
$N_{BP}$ Sized number of parallel batteries \\
$N_{BP}min$ Total minimum number of parallel batteries \\
$N_{BS}$ Sized number of series batteries \\
$N_{BS}min$ Total minimum number of series batteries \\
NOCT Nominal Operating Cell Temperature \\
$N_{PS}$ Sized number of panels in series \\
$N_{PSmin}$ Total number of panels in series \\
$N_{PP}$ Sized number of panels in parallel \\
$N_{PPmin}$ Total umber of panels in parallel \\
$ N_{TP} $ Sized number of panels \\
$N_{TPmin}$ Total minimum number of needed solar panels \\
PV Photovoltaic \\
P Power \\
$P_{in}$ Input power \\
$P_{out}$ Output power \\
$P_{surge}$ Maximum peak power demanded by house for few seconds \\
$q$ Absolute value of the electron's charge ($ -1.60217646\times10^{-19}C $) \\
SAT Satisfiability or Boolean satisfiability problem \\
SAT When returns a failure from the verification tool \\
SMT Satisfiability modulo theories \\
$ SOC $ Battery state of charge \\
STC Standard Test Condition (reference state) \\
$T$ Temperature \\
$ T_{air} $ Ambient temperature \\
$ T_{ref}$ Temperature at STC ($=298.15 K$ or $25^{o}C$) \\
UNSAT Condition when verification tool succeeds (could not reach an error) \\
V Voltage \\
$ V_{AC}  $ Voltage that the inverter supply to the house \\
$V_{c}$ Battery charge voltage  \\
$V_{d}$ Battery discharge voltage \\
$ V_{DC} $ Input voltage to the inverter delivered by the DC source \\
$ V_{in} $ Input voltage \\
$ V_{oc} $ Open-circuit voltage \\
$ V_{oc,ref} $ Reference open-circuit voltage at STC \\
$ V_{out} $ Output voltage \\
$ V_{system} $ DC voltage of the bus \\
$V_{T}$ Thermal voltage ($ V_{T}=k_{B}T/q $) \\
$\Delta T$ Temperature variation ($\Delta T=T-T_{ref} $) \\
$ \mu_{I} $ Short-circuit current temperature coefficient ($A/K$) \\
$ \mu_{V} $ Open-circuit voltage temperature coefficient ($ V/K $) \\
$\eta_{b}$ Battery efficiency \\
$\eta_{c}$ Charge controller efficiency \\
$\eta_{i}$ Inverter efficiency \\
$\eta_{p}$ PV panel efficiency \\