### 2019-07-03 Tex/Old text/main.tex

\documentclass[12pt,twoside]{report}
%\documentclass[12pt,a4paper]{ltxdoc}
%\documentclass[12pt,a4paper]{article}
\usepackage{lmodern}			% Usa a fonte Latin Modern			
\usepackage[T1]{fontenc}		% seleção de códigos de fonte.
\usepackage[utf8]{inputenc}		% determina a codificação utilizada (conversão automática dos acentos)
\usepackage{hyperref}  			% controla a formação do índice
\usepackage{parskip}			% espaçamento entre os parágrafos
\usepackage{microtype} 			% para melhorias de justificação
\usepackage{morefloats}			% permite mais floats
\usepackage{setspace}
%\singlespacing
\onehalfspacing
%\doublespacing
%\setstretch{1.1}

%\documentclass[10pt,a4paper]{article}
%\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{algorithm,algorithmic}
\usepackage{xcolor}
%Alessandro input: makecell package to permite line break inside tables
\usepackage{makecell}
\usepackage[alf]{abntex2cite}	% citacoes

%\documentclass[12pt,twoside]{report}
%\usepackage[utf8]{inputenc}
%\usepackage{graphicx}
\graphicspath{ {images/} }
%\usepackage{caption}
%\usepackage{subcaption}
\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm,bindingoffset=6mm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
%
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

%\usepackage[style=authoryear,sorting=none]{biblatex}
%\addbibresource{references.bib}

\title{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems: Optimal Sizing and Project Validation}
\author{Alessandro Trindade}
\date{31 July 2019}

\begin{document}

\input{titlepage}

\input{2ndcover}
\thispagestyle{empty}
\pagenumbering{gobble}

\input{fichacatalog}
%\thispagestyle{empty}
\newpage

\input{abstract}
\pagestyle{empty}
%\thispagestyle{empty}

\input{resumo}
%\thispagestyle{empty}

%\chapter*{Dedication}
%To mum and dad
%
%\chapter*{Declaration}
%I declare that..
%
\chapter*{Acknowledgements}
(It is not possible to get here without the help of several people. I hope I have not forgotten any and do no wrong).

I would like to thank God (for the gift of life); Solange and Louise (without the patience and support of my family I could not be here); my parents and grandparents; Lucas (for having believed in me and for the patience); the Electricity Department at UFAM (for the support); Professors Feitosa, Marco Cristo and Barreto from IComp (support and advice); IT support team from IComp (Virtual Machine maintenance); secretary people from IComp (for the administrative support); the financial support from FAPEAM, British Council/Newton Fund and Fundação Amazonas Sustentável; Davide Poggio from University of Sheffield for the HOMER Pro license (Research England QR GCRF Institutional Allocation); and Coventry University team (Elena, Kojo, and Ross for the opportunity to use field deployed photovoltaic solar systems).

\pagestyle{empty}

\tableofcontents

\listoffigures

\listoftables
\newpage 
\input{abbreviations}

\chapter{Introduction}
\pagenumbering{arabic}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[RO,LE]{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
\fancyfoot{}
\fancyfoot[LE,RO]{\thepage}
\fancyfoot[CO,CE]{Chapter \thechapter}
\fancyfoot[LO,RE]{A. B. Trindade}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\input{chapters/introduction}

\chapter{Background}
\input{chapters/chapter02}

\chapter{Methodology}
\input{chapters/chapter03}

\chapter{Automated Formal Verification of Stand-Alone Solar PV Systems}
\input{chapters/chapter04}

\chapter{Automated Formal Synthesis Optimization Method for Sizing of Stand-Alone Solar PV Systems}
\input{chapters/chapter05}

\chapter{Conclusions}
\fancyfoot[CO,CE]{Conclusions}
\input{chapters/conclusion}

\appendix
\chapter{List of Publications}
\fancyfoot[CO,CE]{Appendix}
\input{chapters/appendix}

\fancyfoot[CO,CE]{References}
%\printbibliography
\renewcommand\bibname{References}
\bibliography{trindadeThesis}{}

\end{document}