### 2019-07-23 Tex/Old text/chapters/appendix.tex
thesis updated text

%\textbf{List of Publications}
%
Here there are a list of papers that were written during the Thesis elaboration and the status of every submission until the date of this document.

\section{Journals}

Automated Formal Verification of Stand-alone Solar Photovoltaic Systems, 8-May submitted, in review by Elsevier Solar Energy (Impact Factor 4.374, CAPES A1 Interdisciplinary, CAPES A1 Engineering IV ) https://www.journals.elsevier.com/solar-energy

An automated formal synthesis optimization method for sizing of stand-alone solar photovoltaic systems: case studies and comparative, 29-June submitted, Elsevier Applied Solar (Impact Factor 7.9, CAPES A1 Interdisciplinary, CAPES A2 Engineering IV, CAPES B1 Computer Science) https://www.journals.elsevier.com/applied-energy

\section{Congress}

Automated Formal Synthesis of Optimal Sizing for Stand-Alone Solar Photovoltaic Systems - submitted 06-MAY-2019 - New Ideas papers - ASE 2019 The IEEE/ACM Automated Software Engineering (ASE 2019) Conference