### 2019-07-24 Tex/Old text/chapters/appendix.tex
thesis updated text

%\textbf{List of Publications}
%
Here there are a list of papers that were written during the Thesis elaboration and the status of every submission until the date of this document.

\section{Journals}

Trindade, A., Cordeiro, L. \textbf{Automated Formal Verification of Stand-alone Solar Photovoltaic Systems}, submitted: 8-May-2019 , status: in review (2nr round, since 16-July-2019) by Elsevier Solar Energy (Impact Factor 4.374, CAPES A1 Interdisciplinary, CAPES A1 Engineering IV ) \url{https://www.journals.elsevier.com/solar-energy}

Trindade, A., Cordeiro, L. \textbf{An automated formal synthesis optimization method for sizing of stand-alone solar photovoltaic systems: case studies and comparative}, submitted: 29-June-2019, resubmitted: 09-July-2019, status: under review by Elsevier Applied Solar (Impact Factor 7.9, CAPES A1 Interdisciplinary, CAPES A2 Engineering IV, CAPES B1 Computer Science) \url{https://www.journals.elsevier.com/applied-energy}

\section{Congress}

Trindade, A., Cordeiro, L. \textbf{Automated Formal Synthesis of Optimal Sizing for Stand-Alone Solar Photovoltaic Systems}, submitted: 06-MAY-2019, status: under review by The IEEE/ACM Automated Software Engineering (ASE 2019) Conference (New Ideas Papers) \url{https://2019.ase-conferences.org/track/ase-2019-papers}