### 2019-10-26 Tex/Old text/chapters/appendixB.tex

%\textbf{List of Publications}
%
Here we describe the two tools created during the PhD process to implement and validate the two scientific methodologies of the thesis.

\section{Tools}

\section{How to use it}