### 2019-10-28 Tex/Old text/fichacatalog.tex
Thesis update

\thispagestyle{plain}
%\begin{center}
%    \Large
%    \textbf{Automated Verification Applied to Stand-alone Solar Photovoltaic Systems}
%    
%    \vspace{0.4cm}
%    \large
%    Optimal Sizing and Project Validation
%    
%    \vspace{0.4cm}
%    \textbf{Alessandro Trindade}
%    
%    \vspace{0.9cm}
%    \textbf{Ficha Catalográfica elaborada por Suely O. Moraes - CRB 11/365}
%\end{center}
%
%
\thispagestyle{empty}
\null\vfill

\begin{center}
Ficha Catalográfica elaborada por Suely O. Moraes - CRB 11/365
\begin{tabular}{|p{12.5cm}|}%{p{12cm}}
\hline
\begin{small}
\begin{verbatim}
Trindade, Alessandro Bezerra
Automated Verification of Stand-alone Solar 
Photovoltaic Systems: Optimal Sizing and Project Validation. 
/ Alessandro Bezerra Trindade. - Manaus : UFAM, 2019.

121 p.

Tese (Doutorado) - Universidade Federal do
Amazonas. Instituto de Computação. PPG em Informática, 2019.

Inclui bibliografia.

1. Verificação Formal. 2. Model Checking. 3. Sistemas
Solares Fotovoltaicos. I. Título.

006.32 CDD(22.ed.) MEI2008-049
\end{verbatim}
\end{small}
\\ \hline
\end{tabular}
\end{center}

\thispagestyle{plain}