### 2019-10-28 Tex/Old text/chapters/appendix.tex
Added citations
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

%\textbf{List of Publications}
%
Here there are a list of papers that were written during the thesis elaboration and the status of every submission until the date of this document.

\section{Journals}

Trindade, A., Cordeiro, L. \textbf{Automated Formal Verification of Stand-alone Solar Photovoltaic Systems}, submitted: 8 May 2019, status: accepted 26 September 2019; published on-line on 15 October 2019) by Elsevier Solar Energy (Impact Factor 4.674, CAPES A1 Interdisciplinary, CAPES A1 Engineering IV) \url{https://www.sciencedirect.com/science/article/abs/pii/S0038092X1930965X}~\cite{TrindadeCordeiro19}.

Trindade, A., Cordeiro, L. \textbf{Optimal Sizing of Stand-alone Solar PV Systems via Automated Formal Synthesis}, submitted: 28 September 2019, status: under review by IEEE Transactions on Sustainable Energy (Impact Factor 7.65, CAPES A1 Engineering IV), \url{https://www.journals.elsevier.com/applied-energy}~\cite{abs-1909-13139}.

Trindade, A., Degelo, R., Santos Junior, E., Ismail, H., Silva, H., and Cordeiro, L. \textbf{Multi-core model checking and maximum satisfiability applied to hardware-software partitioning}, Online publication date: 15 November 2017, by International Journal of Embedded Systems (IJES), CAPES B2 Computing Engineering index. DOI:   \url {https://doi.org/10.1504/IJES.2017.10008947}~\cite{TrindadeDJISC17}.

\section{Congress}

Trindade, A., Cordeiro, L. \textbf{Automated Formal Synthesis of Optimal Sizing for Stand-Alone Solar Photovoltaic Systems}, submitted: 06 May 2019, status: rejected by The IEEE/ACM Automated Software Engineering (ASE 2019) Conference (New Ideas Papers) in \url{https://2019.ase-conferences.org/track/ase-2019-papers}. This paper is now the base of a new paper, for the 32nd International Conference on Computer-Aided Verification (CAV) 2020 with deadline in 23 January 2020 in \url{http://i-cav.org/2020/call-for-papers/}.