### 2020-01-10 Tex/Old text/chapters/appendixB.tex

Here we describe the two tools created to implement and validate the two scientific methods/techniques of the thesis.

%---------------------------------------
\section{Tools}
%---------------------------------------

During the doctoral process, the techniques from computer science proposed to deal with issues inherent to solar photovoltaic systems became algorithms and programs. These programs became tools, and the tools were used for the purpose of comparison with simulation tools in order to validate the techniques.

One premise of our tools was that they will be written as an ANSI C program, and to follow the requirements established for the International Competition on Software Verification (SV-COMP) \footnote{https://sv-comp.sosy-lab.org/}. This allows us to write a code that can be executed into three different verifiers without adaptation, with the possibility that in the future, other verifiers can also be used.
%
%\textcolor{red}{the second part of this sentence is unclear...} In general, we pursuit to write code that could be verification tool independent, i.e., that can be used in any model checking tool without adaptation, since we aimed to evaluate different.

With this in mind, the code was written without using of '\# include', or '\# define'. Moreover, based on the fact that a compiler library is not used, it was necessary to write in some functions which perform specific calculations depending on the mathematical models adopted (e.g. natural logarithms and exponents).

%---------------------------------------
\subsection{The Automated Verification Tool (for PV System validation)}
\label{sec:automatedverification}
%---------------------------------------

We split the code into four blocks:

\begin{itemize}
\item \textbf{Block 1}: global variables declaration, with weather data from the location (minimum and maximum solar irradiance and temperature, for 24 hours of the day) as array, requirements, and datasheet information pertaining to each item of equipment of the PV system to be verified;

\item \textbf{Block 2}: support functions (e.g. natural logarithms,  exponents);

\item \textbf{Block 3}: PV system specific functions (charge battery, discharge battery, PV panel generation, sizing check);

\item \textbf{Block 4}: main code with charge, discharge battery control, based on PV generation from solar panels and the battery state of charge, committing to deliver the power and energy that the house requires, in  accordance with the restrictions from the sized system.
\end{itemize}

%---------------------------------------
\subsection{The Automated Synthesis Tool (for PV System Optimal Sizing)}
\label{sec:automatedsynthesis}
%---------------------------------------

We split the code into three blocks:

\begin{itemize}
\item \textbf{Block 1}: global variables declaration, with weather data from the location, user requirements, and datasheet information on each item of equipment of the PV system to be verified;

\item \textbf{Block 2}: support functions (start value at the lowest cost for the list of equipment, synthesis phase function to obtain feasible technical solutions with model checking);

\item \textbf{Block 3}: main code with the iterative control of the verify phase.
\end{itemize}

%---------------------------------------
\section{Instructions on How to Use it}
%---------------------------------------

The only block that requires user intervention to input data is Block 1 listed in~\ref{sec:automatedverification} and \ref{sec:automatedsynthesis} from the automated verification and synthesis tools respectively. If the intention is to change the mathematical model adopted, coding must be done in Block 3 of the automated verification tool Block 2 of the automated synthesis tool.

For the automated verification tool, it is necessary to input the following data (editing and modifying the code):

\begin{itemize}
\item Minimum and maximum solar irradiance (average value for the 24 hours of the day) at location;
\item Minimum and maximum temperature (average for every month of the year);
\item Local insolation (average number of sun hours per day at the location);
\item Number of solar panels in series;
\item Number of solar panels in parallel;
\item Number of series-connected cells from each solar panel;
\item Nominal Operating Cell Temperature;
\item Reference solar irradiance;
\item Reference temperature;
\item Short-circuit current temperature coefficient;
\item Open-circuit voltage temperature coefficient;
\item Reference short-circuit current;
\item Reference open-circuit voltage;
\item Reference maximum current, voltage, and power;
\item Minimum MPPT voltage;
\item Operation current;
\item Solar panel efficiency;
\item Panel area in square meters;
\item DC-bus voltage;
\item Individual battery voltage;
\item Battery bank capacity;
\item State of charge limit;
\item Number of batteries in series;
\item Number of batteries in parallel;
\item Battery autonomy;
\item Battery efficiency;
\item Float, absorption, and bulk battery voltages;
\item Charge controller efficiency;
\item Charge controller current;
\item Charge controller maximum voltage;
\item Inverter output voltage;
\item Inverter AC reference power;
\item Inverter maximum (surge) AC power reference;
\item AC voltage (standard voltage from outlet);
\item Power demand from the house;
\item Surge demand from the house;
\item Energy consumption from the house.
\end{itemize}

in the case of the automated synthesis tool, the following data must be added (editing and modifying the code):

\begin{itemize}
\item Minimum and maximum solar irradiance (average value for the 24 hours of the day) at location;
\item Minimum and maximum temperature (average for every month of the year);
\item Local insolation (average number of sun hours per day at the location);
\item DC-bus voltage;
\item Battery autonomy;
\item State of charge limit;
\item Individual battery voltage;
\item Reference solar irradiance;
\item Reference temperature;
\item For every item on the solar PV panel list, as a matrix (area, efficiency, number of series-connected cells, Nominal Operating Cell Temperature, short-circuit current temperature coefficient, open-circuit voltage temperature coefficient, reference short-circuit current, reference open-circuit voltage, reference maximum power, reference maximum current, reference maximum voltage, maximum voltage on NOCT, cost in USD);
\item For every item on the battery list, as a matrix (efficiency, voltage, capacity $C_{20}$), bulk voltage, float voltage, cost in USD);
\item For every item on the charge controller list, as a matrix (efficiency, nominal current, voltage output, minimum MPPT voltage, maximum current, cost in USD);
\item For every item on the inverter list, as a matrix (efficiency, input DC voltage, output AC voltage, reference AC power, reference maximum AC power, cost in USD);
\item AC voltage (standard voltage from outlet);
\item Power demand from the house;
\item Surge demand from the house;
\item Energy consumption from the house.
\end{itemize}

%---------------------------------------
\subsection{Automated Verification with Incremental ESBMC}
%---------------------------------------

Here we describe precisely how to use the code explained in the previous section (\ref{sec:automatedverification}) using the incremental ESBMC with the Z3 SMT solver that presented the best performance during the comparison with other verifiers. However, in \ref{sec:setup} it was shown how to use the code with other verifiers.

ESBMC can be downloaded from its GitHub in \url{https://github.com/esbmc/esbmc}. Installation commands can be found in \url{https://ssvlab.github.io/esbmc/documentation.html}. There is only the Linux version of the verifier.

\bigskip
\fbox{\parbox{\textwidth}{Command-line: \$ esbmc filename.c -\phantom{}-no-bounds-check -\phantom{}-no-pointer-check -\phantom{}-unwind 100 -\phantom{}-smt-during-symex -\phantom{}-smt-symex-guard -\phantom{}-z3.}}
\bigskip

Where:
\begin{itemize}
\item -no-bounds-check: says to the verifier to not perform array bound check (the aim is not to check array bound violation);
\item -\phantom{}-no-pointer-check: says to the verifier to not check if there is pointer violation;
\item -\phantom{}-unwind 100: limits the bound of search (100 in this case);
\item -\phantom{}-smt-during-symex: enables incremental SMT solving;
\item -\phantom{}-smt-symex-guard: calls the solver during symbolic execution;
\item -\phantom{}-z3: uses Z3 as solver.
\end{itemize}

Some results issued by the ESBMC verifier were shown in \ref{sec:ESBMCverification} when conducting the case studies.

%---------------------------------------
\subsection{Automated Synthesis with CPAchecker}
%---------------------------------------

Here we describe precisely how to use the code explained in previous sections (\ref{sec:automatedsynthesis}) using CPAchecker with MathSAT solver, which presented the best performance during the comparison with other verifiers. However, in \ref{sec:synthesissetup} it was shown how to use the code with other verifiers.

The CPAchecker can be downloaded from its web site in \url{https://cpachecker.sosy-lab.org/}. There are versions available for Linux and MS-Windows operational systems.

Install a Java Runtime Environment, which is at least Java 8 compatible (e.g., Oracle JRE, OpenJDK). Cf. the web site \url{http://java.oracle.com/}.

\bigskip
\fbox{\parbox{\textwidth}{The command-line in Linux is: \$ scripts/cpa.sh -heap 64000m -config config/bmc-incremental.properties -spec config/specification/sv-comp-reachability.spc filename.c.}} 
\bigskip

Note that the default heap size is 1200m (or 1.2 GB RAM). However, specifying a significant value with '-heap' is recommended if you have more RAM available. In our case, we adjusted the heap to 64000m (64 GB).

Another important issue is the time-out of the execution process: the parameter 'limits.time.cpu' in the file 'resource-limits.properties' in folder 'config/includes' must be edited according to the desired value. In our case, 'limits.time.cpu = 14400s', i.e. 14400 seconds or 240 minutes or 4 hours.

Some results, reports, and graphics, issued by the CPAchecker verifier were illustrated in \ref{sec:synthesisresults}.